
\subsection{Alternative democracy indices}

To examine the robustness of our specification, we selected a set of alternative measures of democracy to see whether we still observe a significant causal effect of democracy on Covid-19 outcomes. The main measure we have used so far is the Demoocracy Index issued by the Economist Intelligence Unit in 2019, but we will now test the validity of our results by using the following: the Representative Government measure in 2019 by International IDEA, and the Polity Index in 2018 by the Center for Systemic Peace. 

Figure \ref{fig:first-stage-indices} shows the first-stage relationship between the measures of democracy and our main instrument, European settler mortality. As we can observe, the relationship between the indices and settler mortality is negative, which is in accordance with what we would expect. 
    
The results of the 2SLS regression using alternative indices are presented in Table \ref{tab:2sls-indices}. We find that the use of different indices for democracy provide results that are consistent with our previous estimates of a causal effect of democracy on worse performance in the Covid-19 pandemic. In Panel A, we can see that in columns (3), the estimates using the Representative Government Index is statistically significant and point to a positive and large effect of democracy on Covid-19 related deaths per million (419.7 with robust standard error of 226.7). Although the results in (4), (5) and (6) in Panel A are not statistically significant at the 10\% level, we continue to find positive coefficients. In Panel B, where the dependent variable is the change in GDP between 2019 and 2020, the results are even more assuring, as we find statistically significant coefficients for both the Representative Government Index and the Polity Index regardless of whether wee introduce controls or not. More specifically, with the Representative Government Index the estimates are -4.3 without controls and -7.3 with controls; with the Polity index the estimates are -5.4 without controls and -6.0 with controls. Overall, we find that the use of other indices for democracy does not prevent us from concluding that democracy has caused worse performance in the Covid-19 pandemic in both mortality and the economy. 



%\subsection{Alternative controls and weighting}

%The validity of our 2SLS results depends on the assumption that settler mortality in the past has no direct effect on current performance in the pandemic. Although this presumption appears reasonable, here we substantiate it further by directly controlling for many of the variables that could plausibly be correlated with both settler mortality and outcomes in the pandemic, and checking whether the addition of these variables affects our estimates. Overall, we find that our results change fairly little. 
    
%To examine the robustness of our specification, we selected a set of candidates to be additional fundamental determinants and consider a range of specifications.
    
% \subsection{Omitted Variable Bias for IV}


